NoTCP: steganographic tricks to bypass middleboxes

### 2014-07-15 paper/paper.tex
another iteration, Narseo's comments

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
TCP is the de facto transport protocol in the Internet. However it has many limitations that make it ill suited to the modern world of high mobility and multi-homing. Protocol optimizations and filtering performed by middleboxes make extension deployment challenging which has led researchers to explore new ways of exchanging control information between endpoints. In this paper we explore steganographic approaches to smuggle information between endpoints. We design NoTCP -- a mechanism which uses the TCP header fields in combinations that are currently have no defined semantics to hide a small amount of control information from middleboxes and hence bootstrap a control channel. We also present a preliminary study showing how the mechanism works across cellular and WiFi networks.
\end{abstract}

@@

TCP is everywhere in the Internet. The overwhelming majority of bytes are carried in TCP streams~\cite{John:2007dv}, and TCP is used for all types of traffic from bulk download to video streaming to instant messaging. Over the last decade a number of new protocols have been proposed to replace TCP. However many middle boxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like. Consequently few of these protocols have found any leverage.

@@

Instead we believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header. This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.

A practical method of exchanging control information outside of TCP options could not only speed up deployment of existing TCP extensions (\eg MPTCP~\cite{Handley:vj}, TCPCrypt~\cite{Mazieres:uz}), but also help add a generic control stream within TCP. Potential applications range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}. Most importantly, development of new functionality should not be limited by network middleboxes.

This paper makes three main contributions. 
\begin{enumerate}
\item We discuss the possible ways to conceal information in the TCP header and give some examples of the extensions this could enable
\item We describe a measurement application that we have written to run on Android phones.  
\item We present the results of a small scale deployment of our application.
\end{enumerate}

Our hope is to motivate the research community into exploring the full range of possibilities in this space and to encourage more people to install our application so that we can build a comprehensive database of the impact of middleboxes across different network providers.

% Extensibility of the Internet has been not like originally envisioned for a long time now, starting with IP layer and extending the \emph{narrow waist} to TCP or even HTTP at large. The problem is caused by the various middleboxes that aim to improve performance\cite{Kopparty:2002ht,Chakravorty:2003dm}, increase security\cite{Handley:2001vp,Vutukuru:2008fc} and solve address shortage problems. Even congestion control has required multiple extensions since protocol's inception and the large number of proposed congestion control schemes~\cite{5462976} suggests that more will be necessary.

% In an ideal world all infrastructure is forwards compatible and changes together with the needs of the users. Instead, TCP has become the de facto transport protocol of the Internet and development of its extensions has been very challenging. Proposals including MPTCP~\cite{Handley:vj} for multi-homed endpoints and TCPCrypt~\cite{Mazieres:uz} for ubiquitous end-to-end traffic encryption are necessary for today's networks but deployment has been slow and difficult. Previous work on MPTCP overviews the major problems with different types of middleboxes~\cite{Raiciu:2012ur}.

% Although TCP has been built to be extensible via options and reserved bits, all extensions are primarily designed around the limitations of existing network middleboxes\cite{Honda,Guha2005tb,UntoldMiddlebox2011}, but even if options were not filtered, we have run out of the 40-byte TCP option space.

% Problems with the size of option space arise in MPTCP when its options are combined with those of regular TCP and therefore needs to change the semantics of a duplicate ACK~\cite{Handley:vj} and there are efforts to define an alternative control stream to work around the limitation of option space~\cite{Bonaventure:wx}. TCPCrypt uses the Data portion of TCP segments to exchange cryptographic keys~\cite{Mazieres:uz}. An alternative way of looking at the problem is to hide the information from the middleboxes as done by steganographic approaches~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz}.

% We propose \emph{noTCP}, an approach complementary to option negotiation which repurposes TCP header fields beyond what is specified. At the same time it is designed to work across most of deployed network infrastructure. At the core of the idea is modification of TCP header fields in a way that is not interfered with by middleboxes or so that such interference can be safely recovered or ignored. The idea builds up from the work on steganography using TCP~\cite{Frczek:2012dl,Murdoch:2005fz}, but we shift the focus from using sequence numbers and options to other fields as well.

% With the focus on cellular and WiFi networks due to their renown middlebox peculiarities, we make the following contributions in this work:

% \begin{enumerate}
%     \item Review the aspects of the TCP protocol that can potentially be used to extend it (Section~\ref{sec:protocolChanges}).
%     \item Provide an analysis of current mobile (WiFi and cellular) network behavior in the presence of such modifications (Section~\ref{sec:network}).
%     \item Propose a robust method to bootstrap control-stream between two endpoints across middleboxes (Section~\ref{sec:protocol}).
% \end{enumerate}

@@

Without using the options field we only have a few options for control information exchange:
\begin{itemize}
    \item Setting header fields to specific values without setting the corresponding flags to have no meaning to legacy implementations.
    \item Assigning one or more of the reserved field bits.
    \item Defining new semantics for valid header fields.
\end{itemize}

In this section we discuss them and their pros and cons in depth. Fig.~\ref{fig:header} shows TCP header with fields that we propose for reuse and define in Section~\ref{sec:protocol}.

@@

It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics. On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}. Better yet, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and can be redefined.

@@

Only 3 TCP Reserved bis have been left unallocated with the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.

Due to simplicity and non-interference with legacy implementations there are proposals that make use of them. One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets. To avoid triggering duplicate acknowledgments reserved bit is used to mark such packets as duplicate and should be ignored by fast retransmit.

@@

The remaining choice is to encode information within valid header fields. It involves generating packets with specific values for initial sequence numbers, window size or checksums. We posit that it would be an acceptable choice as long as it is deployable in today's network.

Initial sequence numbers must be hard to predict as they provide a measure (however weak) of security and until strong encryption, \eg TCPCrypt, is universally adopted. Adding any information to the field in a way that is easily decoded by the receiver makes it predictable and weakens security  which has already been problematic~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.

@@

To verify current network behavior we have developed an Android application distributed through the \emph{Google Play Store} to collect data from volunteers. It contains a simplified implementation of TCP to generate packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown. It allows us to modify various header fields to our purposes and check the values received by the other endpoint.

The testing setup is:
\begin{itemize}
    \item Android UI to launch tests, report aggregated results to the end-users, send results to the reporting server and manage the tests.
    \item Simplified TCP stack using Raw Unix sockets for full control over the TCP header (\emph{SOCK\_RAW}). %This also meant having to drop RST packets generated by the kernel network stack during the test in response to incoming packets associated with connections it is not aware of - packets are delivered to both the raw socket and the kernel network stack.
    \item Server instrumented to generate specific responses to incoming packets based on header and payload values. We currently use a dedicated test server in a single location and no firewalls or NATs.
\end{itemize}

Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks. The tests were run separately, using previously developed techniques~\cite{Weaver:RHwbx82O}. The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if connection succeeds it is to a proxy on the path. We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers (Section~\ref{sec:portspec})

Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic, whether non-HTTP traffic on HTTP port is filtered, HTTP traffic is normalized (\ie deliberately modified headers are fixed). We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint. Results are presented in Section~\ref{sec:proxies}.

@@

Potential limitations of our methodology include that we do not consider packet splitting and coalescing. Our goal is to exchange control information within packet headers that correct behavior does not depend on how packets are resegmented. To ensure that it does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way. Due to the same reason we do not take congestion control into account either.

Raw sockets operate in parallel to the kernel network stack. We had to suppress reset packets generated by the kernel in response to TCP packets from sessions it is not aware of. We also had to choose local port numbers unused by the kernel. We chose them randomly from the unprivileged port range and relied on other error detection mechanisms to prevent invalid packets from being delivered in the unlikely case of coinciding connection parameters.

@@

@@

@@

We show our collected results in Table~\ref{tab:networks}. Since we were focusing on mobile applications, in our preliminary study we collected results across a few public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries: UK, Finland, Germany and Lithuania.

The results include which networks did not filter any modifications (one WiFi and two cellular nets), and which ones discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers. Only one network provided clients with a global IP (WiFi edu 2).

In traditional NATs~\cite{Egevang:tu} checksum recalculation is very simple: subtract the old header fields from the checksum and add the new ones to minimize the resource use. The networks that deliver packets with invalid checksum only do that as otherwise a packet with invalid checksum would be dropped.

A bigger issue we have discovered is that middleboxes do not always respect the standard and simply drop SYN-ACK packets with payload. We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}.

@@

Finally we tested interaction reserved header bits. We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe different behavior for these: all reserved bits were either allowed through or cleared at both phases. Importantly, we did not observe cases when such packets are dropped.

@@

We repeat our tests on a few ports:
\begin{itemize}
    \item 80, 443 - HTTP and HTTPS
    \item 993 - Secure IMAP port
    \item 5228 - Google cloud messaging port
    \item 6969 - random port number
    \item 8000 - common HTTP proxy port number
\end{itemize}

We only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}. In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped. Our traffic was not encrypted and an increasing proportion of traffic is served over SSL. A commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}, making the value of such proxying questionable.

We confirmed the middlebox to be a ByteMobile proxy. It replaces the original connections with a new one which adds options not present in original packets, changes window size and resets other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}. Such behavior makes deployment of any protocol modifications or extensions very difficult (e.g. we verified that TCPCrypt does not work on this network). Although we did not see more cases of it, the company claims over 160 clients.

In the second case, traffic was only modified when going to ports 80 and 443. As we discuss below, the network optimizes HTTP traffic and since our traffic does not resemble valid HTTP, it is dropped.

@@

Since proxies make new changes very difficult, we checked for non-transparent oes on in our tested networks on a number of popular port numbers:
21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
\hline
    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
    % T-Mobile, part of EE
    Cellular 4*     & HSPA                     & 25, 80, 110, 143 \\ \hline
    % Virgin on T-Mobile
    Cellular 4**    & HSPA                    & 25, 80, 110, 143 \\ \hline
\end{tabular}
\end{center}
}
\vspace{-2mm}
\caption{Non-transparent proxies and proxied port numbers by radio technology}
\label{tab:proxies}
\vspace{-4mm}
\end{table}

Results are summarized in Table~\ref{tab:proxies}. Cellular 2 proxies almost all of our tested ports except for 443 and 993 as observed before. 
Cellular 4 shows non-uniform behavior across the network. It is not obvious why ports commonly used for encrypted communication are proxied over LTE, we only recorded 1 case for HTTPS. The other two cases (* and **) seem related to the operator being formed from two merged operators. These differences are evident in \emph{whois} records and device-reported network name --- \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO), which rent network infrastructure from the mobile network operator.

It is not safe to estimate proxying just from \emph{whois} records: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios. An active test is needed reliably detect proxies.

Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions. While other networks did not interfere with our tests, they do not appear to be completely free of proxies. On Cellular 1 (15 sessions) two sessions show DNS proxies and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80. Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic. We did not have data about Cellular 6, but results discussed above suggest that there are no non-transparent proxies present.

All other networks did not show non-transparent proxy behavior, but the ones who did pose a challenge that is not easily solved. Proxies potentially change every aspect of the original connection and hiding any information within the packet becomes extremely difficult since any value that a proxy does not recognize is reset.

@@

A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections. In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant. The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Fig.~\ref{fig:header}), carrying the desired opcode. For simultaneous open we can use the same mechanism in both directions.

@@

% \vspace{-2mm}
\begin{align*}
Checksum_2 & = Checksum_1 - header_1 + header_2 \\
Checksum_i & = Checksum_{i-1} - header_{i-1} + header_i \\
 ... \\
Checksum_n & = Checksum_1 - header_1 + header_n
\end{align*}

@@

% \vspace{-2mm}
\begin{align*}
Checksum_{target} & = Opcode + header_1 \\
Checksum_n = Opcode_C & = Opcode + header_n
\end{align*}

@@

We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream. We extend the three-way handshake with fields shown in Fig.~\ref{fig:header}:

@@

An example handshake is shown in Fig.~\ref{fig:handshake}. In this case a client (active opener) sends a SYN packet as show above. The packet traverses a middlebox which does not remove the opcode and reaches the Server (passive opener). The Server replies with SYN-ACK with U\_Opcode which is discarded by the middlebox. It then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, which reaches the Client. The client recovers and verifies the received opcode. Both endpoints are now assumed to understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.

Putting control stream into packet payload suffers from much the same issues as other proposals to extend TCP option space~\cite{Ramaiah:2012wa}. The premise of our work is that middleboxes should not be aware of the new control streams in the first place and therefore not interfere with them.

We witnessed a case of resegmentation where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload. It is not an issue with TLV-encoded control stream in the payload, because it will be delivered to the receiver first. The mechanism of \emph{Checksum Correction} would, however, be compromised for the first segment. If the passive opener receives data during or immediately after the handshake but can not decode that as valid TLV-encoded control information, it treats that as application data and falls back to vanilla TCP.

The focus of our work is to show the possibility of exchanging control information within the packet header. We therefore leave refinement of the full control channel for future work.

@@

TCP implements an \emph{urgent mechanism} at its core that allows the sending user to stimulate the receiving user to accept some \emph{urgent data}. The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data. Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and in general it is recommended against the use of the mechanism~\cite{Gont:2011vi}.

@@

Generic control stream for MPTCP~\cite{Bonaventure:wx} can be reused as well. It proposes the mapping control stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams. The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be compatible with the new specification.

%More generally SCTP uses \emph{data chunks} to differentiate between different subflows. Each packet can contain multiple TLV-encoded chunks to separate user data and control information. Despite its challenged deployment in the Internet, SCTP over DTLS is being used as the basis of WebRTC data protocol~\cite{Tuexen:wv} and we suggest that such chunking could be used on top of \emph{noTCP} once control channel has been bootstrapped.

@@

The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, is it deployable and is it forwards-compatible? We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.

@@

In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define which header fields it interferes with and how. Examples include extending sequence numbers and using stronger message integrity checks~\cite{Mazieres:uz}, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple checksum or window size modifications.

>>>>

% end author
\maketitle
% Use the following at camera-ready time to suppress page numbers.
% Comment it out when you first submit the paper for review.
%\thispagestyle{empty}
\begin{abstract}
TCP is the de facto transport protocol in the Internet. However it has many limitations that make it ill suited to the modern world of high mobility and multi-homing. Protocol optimizations and filtering performed by middleboxes make extension deployment challenging which has led researchers to explore new ways of exchanging control information between endpoints. In this paper we explore steganographic approaches to smuggle information between endpoints to avoid middlebox interference. We design NoTCP -- a mechanism which uses the TCP header fields in combinations that are currently have no defined semantics to hide a small amount of control information from middleboxes and hence bootstrap a control channel. We also present a preliminary study showing how the mechanism works across cellular and WiFi networks.
\end{abstract}

@@

TCP is everywhere in the Internet and it carries the overwhelming majority of bytes~\cite{John:2007dv}. TCP is used for all types of traffic from bulk download to video streaming to instant messaging. Over the last decade a number of new protocols have been proposed to replace TCP. However many middle boxes will actively block or interfere with traffic that does not conform to their view of what TCP should look like. Consequently few of these protocols have found any leverage.

Modern networks necessitate new functionality, such as Multipath Connectivity~\cite{Handley:vj} for multi-homed devices and ubiquitous traffic encryption~\cite{Mazieres:uz}). The applications of exchanging control information between endpoints range from new secure extensions~\cite{Rescorla:R080FoGB} to exchanging information about endpoints for connection parameter tuning (\eg initial congestion window~\cite{Dukkipati:2010hs}) to designing better coordination for long-lived connections to save mobile device energy~\cite{Aucinas:2013uk}.

@@

We believe that the solution is to work round the middle boxes by using steganographic techniques~\cite{Frczek:2012dl,Zielinska:2014fn,Rowland:1997vq,Murdoch:2005fz} to conceal control information within the TCP header. This could be as simple as using one of the currently undefined flags to toggle the meaning of existing fields, through to using the urgent pointer without setting the urgent flag or even hashing information into the initial sequence number in much the way that SYN Cookies~\cite{Eddy:2007to} do currently.

This paper makes three main contributions. 
\begin{enumerate}
\item We discuss the possible ways to conceal information in the TCP header and give some examples of the extensions this could enable
\item We describe a measurement application that we have written to run on Android phones for a large-scale study.  
\item We present the results of an initial small scale deployment of our application.
\end{enumerate}

We use our initial study to design a way of exchanging a small amount of control information successfully across all of our tested networks and show a potential mechanism to use it for bootstrapping a full control stream currently missing from TCP. We hope that our successful results will motivate the research community into exploring the full range of possibilities in this space.

@@

Figure~\ref{fig:header} shows TCP header with fields that we propose for reuse and define later in the paper. However, the structure of TCP header is very simple and gives are few choices for hiding extra control information:
\begin{itemize}
    \item Header fields without corresponding flags to have no meaning to legacy implementations.
    \item Assigning one or more of the reserved field bits.
    \item Defining new semantics for valid header fields.
\end{itemize}

@@

The general recommendation is to not use It is recommended against the use of the urgent mechanism~\cite{Gont:2011vi}, therefore it could be reused for other purposes if both endpoints understand the new semantics. On the other hand, it is common practice by Network Intrusion Detection Systems (NIDS) on enterprise networks to reset the field as it makes it difficult to track application-layer data~\cite{seolma}. Better yet, the field becomes \emph{Non-Urgent}~\cite{Kuhlewind:2014vd} when URG is not set -- it has no meaning for legacy implementations and can be redefined.

@@

Only 3 TCP Reserved bits have not been allocated since the three ECN bits (ECN, CWR and experimental NS)~\cite{Floyd:up,Ely:uc} that are still not universally implemented~\cite{Kuhlewind:2013hr}.

Due to simplicity and non-interference with legacy implementations there are proposals that make use of them. One suggestion for reducing Web latency~\cite{Flach:2013uy} by mitigating the effects of a small number of lost packets for short connections is to send multiple copies of packets. The solution avoids triggering fast retransmit by such duplicate acknowledgments by marking them as duplicates in a reserved flag.

@@

The remaining choice is to encode information within valid header fields. It involves generating packets with specific values for initial sequence numbers, window size or checksums. Arguably it is an acceptable choice if deployable in today's networks.

Initial sequence numbers must be hard to predict as they provide a measure of security and until strong encryption. Stronger security measures such as TCPCrypt are better suited for this purpose, but are not universally adopted. Adding information to the field to be easily decoded by the receiver makes it predictable and weakens already problematic security~\cite{Qian:2012bj,Bellovin:uz,Qian:2012wb}.

@@

To verify current network behavior we have developed an Android application distributed through the \emph{Google Play Store} to collect data from volunteers. It contains a simplified implementation of TCP to generate packet sequences that appear valid to network middleboxes: three-way handshake, data exchange and connection teardown. It allows us to modify various header fields to our purposes and check the values received by the other endpoint. We control both endpoints of each connection, therefore we can detect changes from sent values received by the other endpoint.

The testing setup is:
\begin{itemize}
    \item Android application with a simplified TCP stack using Raw Unix sockets for full control over the TCP header (\emph{SOCK\_RAW}). It manages the tests and sends results back to our servers. 
    \item Server instrumented to generate specific responses to incoming packets based on header and payload values. We currently use a dedicated test server in a single location and no firewalls or NATs.
\end{itemize}

Non-transparent proxies are a major problem for extension deployment, therefore we used another well know and widely deployed tool to detect proxies on our tested networks. We implemented previously developed techniques~\cite{Weaver:RHwbx82O} and run the test separately. The primary one was non-responsive server test: the server is instrumented not to respond to SYN, therefore if connection succeeds it is to a proxy on the path. We repeated these tests on different destination ports and found that proxying can be specific to certain port numbers.

%Transparency of proxies may also depend on higher-level protocols, therefore we checked for HTTP proxies using techniques proposed in the work, including checking content and header changes, different IP addresses seen for HTTP or non-HTTP traffic and whether non-HTTP traffic on HTTP port is filtered.

@@

Potential limitations of our methodology include that we do not consider packet splitting and coalescing. Our goal is to exchange control information within packet headers to avoid dealing with it altogether. To ensure that splitting does not happen during our tests we do not send more than a few bytes of data and the exchange only consists of a single message sent each way. Due to the same reason we do not take congestion control into account either.

Raw sockets operate in parallel to the kernel network stack. We had to suppress reset packets generated by the kernel in response to TCP packets from sessions it is not aware of. We also had to choose local port numbers unused by the kernel. We chose them randomly from the unprivileged port range and relied on other error detection mechanisms preventing invalid packet delivery to the applications.

@@

\vspace{-3mm}

@@

\vspace{-2mm}

@@

We show our collected results in Table~\ref{tab:networks}. Since we were focusing on mobile applications, in our preliminary study we collected results across public, residential, university and enterprise WiFi networks as well as 6 cellular networks in 4 different countries.

The results include networks that did not filter any modifications (only one WiFi and two cellular nets), and ones that discarded packets with invalid checksums, normalized particular fields - either dropped packets or reset to different values - and whether NATs performed only port number remapping or also changed sequence numbers. Only one network provided clients with a global IP (WiFi edu 2).

Traditional NATs~\cite{Egevang:tu} perform a simple checksum recalculation: subtract the old header fields from the checksum and add the new ones. The networks that deliver packets with invalid checksum only do that. Otherwise they would detect and drop an invalid packet.

A bigger issue we have discovered is that middleboxes do not always respect the standard and drop SYN-ACK packets with payload. We did not see it on cellular networks, but most WiFi (4/7) behaved in such way even though the standard allows handshake packets to contain payload~\cite{Postel:3EDyoxP_,Chu:2011tn}.

@@

Finally we tested interaction reserved header bits. We separated our tests to handshake and data exchange parts as well as set each bit individually, however we did not observe any differences. All networks appear to pass reserved bits on non-proxied connections.

@@

To differentiate between general behavior of particular network and application-specific optimizations (\eg HTTP acceleration) we repeat our tests on a few application-specific ports as well as a random one:
\begin{itemize}
    \item 80, 443 - HTTP and HTTPS
    \item 993 - Secure IMAP port
    \item 5228 - Google cloud messaging port
    \item 6969 - random port number
    \item 8000 - common HTTP proxy port number
\end{itemize}

We only observed port-specific behavior on two networks: \emph{Cellular 2} and \emph{Cellular 4}. In the first case, traffic travels over SSL ports (443, 993) unmodified except for packets with invalid checksum being dropped. Our traffic was not encrypted. An increasing proportion of traffic is served over SSL, making the value of such proxies questionable: a commercially deployed HTTP accelerator reports the proportion of HTTPS requests to be close to that of HTTP\footnote{\url{http://db.awazza.com/users/global/}}.

We confirmed the middlebox on Cellular 2 to be a ByteMobile proxy. It replaces the original connections with new ones, adding options not present in original packets, changing window size and resetting other fields to new values to optimize TCP for cellular networks~\cite{Ha:2006td}. Such behavior makes deployment of any protocol modifications or extensions difficult -- we verified that \eg TCPCrypt does not work on this network. We do not currently have an estimate of how widespread ByteMobile proxies are as we only observed them on one network.

In the second case, traffic was only modified when going to ports 80 and 443. The network likely optimizes HTTP traffic and drops any non-HTTP traffic. In general port-specific behavior appears related to proxies.

@@

Proxies potentially change every aspect of the original connection, rewrite or filter header values they do not recognize and hiding any information within the packet becomes difficult. We checked for them in our tested networks on a number of popular port numbers:
21, 22, 25, 80, 110, 135, 139, 143, 161, 443, 445, 465, 585, 587, 993, 995, 1194, 1723, 5060, 6881, 9001.

\begin{table}[t!]
{\small
\begin{center}
\begin{tabular}{| l | >{\centering\arraybackslash}m{1.8cm} | >{\centering\arraybackslash}m{3.7cm} |}
\hline
    \textbf{Net ID} & \textbf{Radio Technology} & \textbf{Proxied Ports} \\ \hline \hline
    Cellular 2      & HSPA                      & 21, 22, 25, 80, 110, 135, 143, 161, 465, 585, 587, 995, 1194, 1723, 5060, 6881, 9001 \\ \hline
    Cellular 4      & GPRS                      & 80, 110, 143 \\ \hline
    Cellular 4      & HSPA                      & 80, 110, 143 \\ \hline
    Cellular 4      & LTE                       & 80, 110, 143, 443, 993, 995 \\ \hline
    % T-Mobile, part of EE
    Cellular 4*     & HSPA                     & 25, 80, 110, 143 \\ \hline
    % Virgin on T-Mobile
    Cellular 4**    & HSPA                    & 25, 80, 110, 143 \\ \hline
\end{tabular}
\end{center}
}
\vspace{-2mm}
\caption{Non-transparent proxies and their proxied port numbers by radio technology}
\label{tab:proxies}
\vspace{-4mm}
\end{table}

We summarize results in Table~\ref{tab:proxies}. 3G and 4G networks use different IP core networks, therefore we split the results by radio technology. We did not detect proxies on Cellular 1 and Cellular 3 networks and did not have data about Cellular 6, but results of our TCP testsuite did not indicate any proxies. Cellular 2 proxies almost all of our tested ports except for 443 and 993.

Cellular 4 shows non-uniform behavior across the network. It is not obvious why the network proxies connections to ports commonly used for encrypted communications. Currently we have only recorded 1 such case for HTTPS. The differences between the other two cases (* and **) seem related to forming the operator from two merged ones. Their infrastructure may have not been completely consolidated: \emph{whois} records and device-reported network names are different. Case \textbf{*} shows the old operator name and \textbf{**} reports the name of a virtual network operator (MVNO). MVNOs buy network services from operators at wholesale rates and typically have little infrastructure of their own, and expect to see the same behavior as that of the underlying network.

Estimating proxying from \emph{whois} records alone is not safe: they do not necessarily match the operator names and we saw the same address ranges used with all device-reported network names for HSPA and LTE radios. Reliable detection needs active testing, potentially combined with caching for repeated connections.

%Proxying is protocol-dependent in some cases: looking at port 80, Cellular 4\*\* modified HTTP headers, performed transcoding and reordered header fields, but only modified non-HTTP traffic for one of the two sessions. While some networks did not interfere with our tests, they do not appear to be completely free of proxies. On Cellular 1 (15 sessions) two sessions show DNS proxies (over UDP) and on Cellular 5 two sessions (out of 47) had non-HTTP traffic filtered on port 80. Finally, Cellular 2 and 4 both reorder HTTP headers and perform transcoding, but Cellular 4 does not always allow non-HTTP traffic.

@@

A surprising result in our tests is that the \emph{Acknowledgment Number} field without the ACK flag set is passed on all non-proxied connections. In addition, it is a 32-bit value and we can allocate part of it (\eg 8 or 16 bits) to signal that the rest of the field is significant. The exact allocation is subject to further discussion, but we suggest using the first 16 bits as \emph{Extension Space} and the others as $Opcode$ (Figure~\ref{fig:header}), carrying the desired opcode. For simultaneous open we can use the same mechanism in both directions.

@@

\vspace{-2mm}
\begin{align*}
Checksum_2 & = Checksum_1 - header_1 + header_2 \\
Checksum_i & = Checksum_{i-1} - header_{i-1} + header_i \\
 ... \\
Checksum_n & = Checksum_1 - header_1 + header_n
\end{align*}

@@

\vspace{-2mm}
\begin{align*}
Checksum_{target} & = Opcode + header_1 \\
Checksum_n = Opcode_C & = Opcode + header_n
\end{align*}

@@

We use the embedded opcodes in the handshake to exchange sufficient control information to bootstrap a complete control stream. We extend the three-way handshake with fields shown in Figure~\ref{fig:header}:

@@

Figure~\ref{fig:handshake} shows an example handshake. In this case a client (active opener) sends a SYN packet as show above. The packet traverses a middlebox that does not remove the opcode and reaches the Server (passive opener). The Server replies with SYN-ACK with U\_Opcode, but the middlebox discards it. The server then retransmits SYN-ACK with $Opcode_C$ and Checksum Correction bits, delivered to the Client. The client recovers and verifies the received opcode. Hence both endpoints understand the meaning of the opcode, therefore the client replies with an ACK with control information acknowledging reception of the $Opcode$ from the Server.

We witnessed a case of packet splitting where an ACK segment is split to one only acknowledging the received data and another other one with the actual payload. We did not observe a case where a middlebox both passes payload in SYN-ACK packet and splits ACK segments. If the handshake succeeds we need to take a little extra care with the fallback mechanism -- we need to wait for an ACK segment with payload before falling back to vanilla TCP. If the passive opener receives data it can not decode as valid TLV-encoded control information, it falls back to vanilla TCP.

Putting control stream into packet payload suffers from much the same issues as other proposals to extend TCP option space~\cite{Ramaiah:2012wa}. The premise of our work is that middleboxes should not be aware of the new control streams in the first place and therefore not interfere with them. Currently we focus on exchanging control information within the packet header and leave refinement of the full control channel for future work.

\subsection{Control channel bitrate}

In progress...

@@

TCP implements an \emph{urgent mechanism} at its core that allows the sending user to stimulate the receiving user to accept some \emph{urgent data}. The mechanism is often quoted as providing ``out-of-bound'' data delivery even though it is explicitly not designed as a mechanism for sending such data. Furthermore, there are ambiguities regarding the semantics of the urgent pointer, Network Intrusion Detection Systems (NIDS) tend to clear the URG flag and pointer and the general recommendation is against the use of the mechanism~\cite{Gont:2011vi}.

@@

Generic control stream for MPTCP~\cite{Bonaventure:wx} is another alternative. It proposes the mapping control stream into a separate sequence number space and exchange control data over established subflows, only modifying the MPTCP DSS option to differentiate control and data streams. The underlying assumption, however, is that MPTCP is already deployed and that existing deployments will be compatible with the new specification.

@@

The main points of discussion on \emph{NoTCP} are the same as with all TCP extensions: is it necessary, deployable and forwards-compatible? We have discussed the first two throughout the paper, but the biggest challenge for forwards-compatibility is NoTCP's overloading of header fields instead of using options.

@@

In the long term and beyond our example control information exchange we believe that each protocol-relate standards (TCP extensions in this case) should explicitly define all header fields it interferes with and how. Examples include extending sequence numbers and using stronger message integrity checks~\cite{Mazieres:uz}, changing semantics of duplicate ACKs~\cite{Handley:vj,Flach:2013uy} or even simple checksum or window size modifications.
